Sequence Labeling for Gait Analysis using LSTM

### 2015-05-09 paper/paper.tex
Minor fixes

In our opinion, the feed-forward neural network approach has several
disadvantages.
In addition to heavy preprocessing consisting of feature extraction and PCA, the
data has to be split up into windows of variable length.
This in turn requires that the foot displacement be measured accurately from
the data which is non-trivial.
Moreover, the network architecture seems to be too simplistic to effectively
capture the long-range dependencies that exist between the data samples.

@@

Gait event detection can be seen as a form of sequence labeling (sometimes also
called segment classification), i.e., the task of assigning a sequence of labels
(swing phase vs. stance phase) to a sequence of input data (the motion capture
signals).
The success of a sequence labeling system depends crucially on its ability to
incorporate context information from both sides of the segment in question.
%
One way to incorporate this context information is to use feed-forward or
recurrenct neural networks with a sliding windows (as in the case of
\cite{Miller2009}).
However, as the range of useful context and thus the required window size is in
general not known in practice, this approach is only of limited usefulness
\cite{Graves2012}.

@@

\section{Data}
\label{sec:Data}
Our training and testing was conducted on data from 10 healthy subjects walking at three different speeds for 50 seconds. Each sequence of 50 seconds will be called a trial from now on. This totals 1500 seconds of recorded data, which roughly correspond to 1000 gait cycles.  For each subject the motion-capture data (input) is 3D marker position of 18 different markers placed in the same anatomical positions for each subject, sampled at 100Hz. The event labels (target) were computed from ground force reaction data sampled at 1kHz.

@@

>>>>

The feed-forward neural network approach has several disadvantages.
In addition to heavy preprocessing consisting of feature extraction and PCA, the
data has to be split up into windows of variable length.
This in turn requires that the foot displacement be measured accurately from
the data which is non-trivial.
Moreover, the network architecture seems to be too simplistic to effectively
capture the long-range dependencies that exist between the data samples.

@@

Gait event detection can be seen as a form of sequence labeling (sometimes also
called segment classification), i.e., the task of assigning a sequence of labels
(swing phase vs. stance phase) to a sequence of input (motion) data.
The success of a sequence labeling system hinges on its ability to
incorporate context information from both sides of the segment in question.
%
One way to incorporate this context information is to use feed-forward or
recurrenct neural networks with a sliding windows (as in the case of
\cite{Miller2009}).
However, as the range of useful context and thus the required window size is in
general not known in practice, this approach is only of limited usefulness
\cite{Graves2012}.

@@

\section{Data}
\label{sec:Data}
Our training and testing was conducted on data from 10 healthy subjects walking 
on a treadmill at three different speeds for 50 seconds. Each sequence of 50 seconds will be called a trial from now on. This totals 1500 seconds of recorded data, which roughly correspond to 1000 gait cycles.  For each subject, 3D motion capture data was collected using 18 reflective markers placed in the same anatomical positions for each subject, sampled at 100Hz. The event labels (target) were computed from ground force reaction data sampled at 1kHz.

@@

Overall, the results we obtained are satisfying and we expect to polish them a little further before submitting them for publication in a gait research journal.
